---
title: "R Notebook"
output: html_notebook
---


```{r setup, include=FALSE}
#install.packages('brazilmaps')
#install.packages("readxl")
knitr::opts_chunk$set(echo = TRUE)
#library(brazilmaps)
#library(ggplot2)
library(readxl)
library(rjags)
library(reshape2)
library(HDInterval)
```

## Import data
```{r}
co4<-read.csv('../Data/co.meso.v1.csv')
co4<-co4[co4$mesoregion!='NA_low',]

```

##Which age subset do we want to analyze?
```{r}
co4<-co4[co4$agec %in% c(2,3),] #Subsets the datasets. In Colombia age groups 1 and 2 are 0-23m. Can't separate out <2m

```

```{r}
co4<-co4[co4$Year<=2015,]
co4<-co4[!is.na(co4$mesoregion),]
```


```{r}
plot(co4$PCV2.coverage[co4$Year==2013], co4$PCV3.coverage[co4$Year==2013], xlim=c(50, 140), ylim=c(50,140))
abline(a=0, b=1)
```

```{r}
lst <- strsplit(as.character(co4$mesoregion), split="_", fixed=T)
lst<-sapply(lst, function(x) x[2])
co4$gdp.level<-lst
```


Reshape data for JAGS
```{r}
#reshape the data for JAGS
co4.m<-melt(co4[,c('mesoregion','Year','agec','j12_j18','ach.noj',"e40_46",'r95_99', 'a00_b99','PCV3.coverage' )], id.vars=c('mesoregion','Year','agec'))

#Filter out sparse regions with <1 death/year
filter1<-dcast(co4.m[co4.m$variable=='j12_j18',],mesoregion~Year, sum)
keep <- which(apply(filter1[,-1],1, mean) > 10)
mes.keep <-filter1$mesoregion[keep]

co4.m <- co4.m[co4.m$mesoregion %in% mes.keep,]
co4.m$mesoregion <- factor(co4.m$mesoregion)

#Sums over whichever age groups are in the data
j12.c<-dcast(co4.m[co4.m$variable=='j12_j18',],mesoregion~Year, sum)
j12.c<-j12.c[,-1]
e40.c<-dcast(co4.m[co4.m$variable=='e40_46',],mesoregion~Year, sum)
e40.c<-e40.c[,-1]
r95.c<-dcast(co4.m[co4.m$variable=='r95_99',],mesoregion~Year, sum)
r95.c<-r95.c[,-1]
ach.c<-dcast(co4.m[co4.m$variable=='ach.noj',],mesoregion~Year, sum)
ach.c<-ach.c[,-1, drop=F]
log.ach.c<-log(ach.c)

a00.c<-dcast(co4.m[co4.m$variable=='a00_b99',],mesoregion~Year, sum)
a00.c<-a00.c[,-1]

#Lower middle, upper middle, upper based on USD, multiply by 1.659 to reais basd on Jan 2011
co4$gdp.cat<-NA
co4$gdp.cat[co4$gdp.level=='low']<-1
co4$gdp.cat[co4$gdp.level=='med']<-2
co4$gdp.cat[co4$gdp.level=='high']<-3

co4.m.gdp<-melt(co4[co4$mesoregion %in% mes.keep,c('mesoregion','gdp.cat' )], id.vars=c('mesoregion'))
co4.m.gdp$mesoregion <- factor(co4.m.gdp$mesoregion)
gdp.c<-dcast(co4.m.gdp[co4.m.gdp$variable=='gdp.cat',],mesoregion~., max, na.rm=T)
gdp.c<-gdp.c[,-1]
gdp.med<-gdp.c
gdp.med[]<-0
gdp.med[gdp.c==2 ]<-1
gdp.hi<-gdp.c
gdp.hi[]<-0
gdp.hi[gdp.c==3]<-1

vax.c<-dcast(co4.m[co4.m$variable=='PCV3.coverage',],mesoregion~Year, mean)
vax.c<-as.matrix(vax.c[,-1])
vax.c<-vax.c/100
#vax.c<-apply(vax.c,2, function(x) x-(mean(x))) #center coverage on 0 for each year
```



#Check data
```{r}
j12.c.sum<-aggregate(j12.c, by=list('hdi.cat'=gdp.c), FUN=sum)
ach.c.sum<-aggregate(ach.c, by=list('hdi.cat'=gdp.c), FUN=sum)
e40.c.sum<-aggregate(e40.c, by=list('hdi.cat'=gdp.c), FUN=sum)
r95.c.sum<-aggregate(r95.c, by=list('hdi.cat'=gdp.c), FUN=sum)

matplot(t(j12.c.sum[,-1])/ t(ach.c.sum[,-1]), type='l' )
abline(v=5.5)

matplot(t(j12.c.sum[,-1])/ t(e40.c.sum[,-1]), type='l' )
abline(v=5.5)

matplot(t(j12.c.sum[,-1]) , type='l')
abline(v=5.5)

matplot(t(r95.c.sum[,-1]) , type='l')
abline(v=5.5)

matplot(t(j12.c.sum[,-1])/ t(r95.c.sum[,-1]), type='l' )
abline(v=5.5)
```

```{r, fig.width=3, fig.height=3}
scale1.j12 <- apply(j12.c.sum[,-1],1, function (x) x/x['2010'])
matplot(2005:2015,scale1.j12, type='l', col=c('red','black','blue'), bty='l', ylab='Deaths relative to 2010', xlab='')
abline(h=1, lty=2, col='gray')
```

Ave coverage
```{r}

round(sum(vax.c[,'2015']*j12.c$`2010`/sum(j12.c$`2010`, na.rm=T), na.rm=T)*100,1)
```



```{r}
#plot(gdp.c,vax.c[,11])
```


```{r}
library(lme4)
ds.test <- co4
ds.test$scale.cov3 <- ds.test$PCV3.coverage/100

mod1<-glmer(j12_j18~  #as.factor(Year) +
              scale(log(ach.noj+0.5)) +scale.cov3 +scale(log(a00_b99+0.5)) +(scale.cov3|mesoregion), data=ds.test, family='poisson',control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=100000))
)
summary(mod1)
```

Draft of model
```{r}
model_string<-"

model{

for(i in 1:subnat){

   for(t in 1:time){

      y[i,t] ~ dpois(lambda[i,t])
      log(lambda[i,t]) <- x[i,t,]%*%beta[i,] + 
                          vax[i,t]%*%theta[i] +
                          disp[i,t]

      log.rr.vax[i,t] <- vax[i,t]%*%theta[i]
   }
   
   for(j in 1:p_x){
      beta[i,j] ~ dnorm(mu_beta[j], inv_var_beta[j])
   }
      
   theta[i] <-z[i,]%*%gamma  + eta[i]
   eta[i] ~ dnorm(0, inv_var_eta)
}

for(i in 1:subnat){
    disp[i,1]~dnorm(0,tau.disp)
  for(t in 2:time){
    disp[i,t]~dnorm( disp[i,t-1]*rho,tau.disp)
  }
}

rho ~ dunif(0,1)

#https://stackoverflow.com/questions/47135726/error-slicer-stuck-at-value-with-infinite-density-running-binomial-beta-model
# for(i in 1:subnat){
#  rho[i] ~ dbeta((a+0.01),(b+0.01))T(0.001,0.999)
#  tau.disp[i] ~ dgamma(c, d)
# }
# 
#   a ~ dgamma(0.01,0.01)
#   b ~ dgamma(0.01,0.01)
#   c ~ dgamma(0.01,0.01)
#   d ~ dgamma(0.01,0.01)

for (j in 1:p_x) {
   mu_beta[j] ~ dnorm(0, 0.0001)
   inv_var_beta[j] ~ dgamma(0.01,0.01)
}

  gamma[1] ~ dnorm(0, 0.0001) #reference
  gamma[2] ~ dnorm(0, 0.0001)
 # gamma[3] <- step(gamma[2]-gamma[1])*(gamma[2] + delta) +
 #             step(gamma[1]-gamma[2])*(gamma[2] - delta)
  gamma[3] ~dnorm(0, 0.0001)

  delta ~dgamma(0.01,0.01)
  
  ave.effect.pov[1] <- gamma[1] 
  ave.effect.pov[2] <- gamma[1] + gamma[2]
  ave.effect.pov[3] <- gamma[1] + gamma[3]

    inv_var_eta~dgamma(0.01,0.01)
  mu.eta ~ dnorm(0, 1e-4)

  tau.disp~dgamma(0.01,0.01)
  
}
"
```

Call model
```{r}
  ##############################################################
  #Model Fitting
  ##############################################################
  inits1=list(".RNG.seed"=c(123), ".RNG.name"='base::Wichmann-Hill')
  inits2=list(".RNG.seed"=c(456), ".RNG.name"='base::Wichmann-Hill')
  inits3=list(".RNG.seed"=c(789), ".RNG.name"='base::Wichmann-Hill')
  burnN=10000
  sampleN=100000
  ##############################################
  #Model Organization
  ##############################################
  
  x<-array(1,
           c(nrow(j12.c),
             ncol(j12.c),
             2))
  x[,,2]<-as.matrix(log.ach.c)
 # x[,,3]<-as.matrix(log(a00.c+0.5))
#  x[,,4]<-as.matrix(log(r95.c+0.5))

    
  z<-matrix(1,
            nrow=nrow(j12.c),
            ncol=3)
  z[,2]<-gdp.med
  z[,3]<-gdp.hi


  #missing values of vax (for testing purposes only)
  vax.c[is.na(vax.c) == 1]<-0
  
  vax.c.mat <- as.matrix(vax.c)
  mean.vax.c <- mean(vax.c.mat)
  sd.vax.c <- sd(vax.c.mat)
  vax.c.scale <- (vax.c.mat )/sd.vax.c
  
```

```{r, eval=T}
  model_spec<-textConnection(model_string)
  model_jags<-jags.model(model_spec, 
                         inits=list(inits1),
                         data=list('y' = j12.c,
                                   'subnat' = nrow(j12.c),
                                   'time' = ncol(j12.c), 
                                   'x' = x,
                                   'vax' = vax.c.scale,
                                   'p_x' = dim(x)[3],
                                   'p_z' = ncol(z),
                                   'z' = z
                                   ),
                         n.adapt=burnN, 
                         n.chains=1)
  params<-c('beta','theta','gamma','log.rr.vax', 'ave.effect.pov', 'lambda')
  
  ##############################################
  #Posterior Sampling
  ##############################################
  posterior_samples<-coda.samples(model_jags, 
                                  params, 
                                  n.iter=sampleN)
  posterior_samples.all<-do.call(rbind,posterior_samples)
  post_means<-apply(posterior_samples.all, 2, median)
  sample.labs<-names(post_means)
  ci<-t(hdi(posterior_samples.all, credMass = 0.95))
  row.names(ci)<-sample.labs
  names(post_means)<-sample.labs
```

```{r, eval=F, fig.width=6, fig.height=10}
plot(posterior_samples, ask=T)
```

## Calculate DIC
```{r}
#DIC
y<-as.vector(as.matrix(j12.c))
    reg.mean.fitted<- t(posterior_samples.all[,grep('lambda', dimnames(posterior_samples.all)[[2]] )])
    pred.count<-reg.mean.fitted #lambda
    pred.count.mean<-apply(pred.count,1,mean)
    log.like.func<-function(x1){
    neg_two_loglike_poisson<- -2*sum(dpois(as.vector(y), 
                                             lambda = x1, 
                                             log = TRUE))
    }
    log.lik.mat<-apply(pred.count,2,log.like.func) #Object of length D, with -2LL estimates
      #Calculate the mean of the fitted values. Prd.count mean, is a vector of length N,
    #And use this to calculate neg_two_loglike_poisson_mean
    neg_two_loglike_poisson_mean<- -2*sum(dpois(as.vector(y),
                                                lambda = pred.count.mean,
                                                log = TRUE))
    DIC<- 2*(mean(log.lik.mat)) -   neg_two_loglike_poisson_mean
    p_D<-  mean(log.lik.mat)  -   neg_two_loglike_poisson_mean #number of parameters
    DIC
    p_D
```
Gamma2 and Gamma3 tells us the effect of SES
```{r}
beta1<-post_means[grep("^beta.*,1]", names(post_means))]
beta1.ci<-ci[grep("^beta.*,1]", dimnames(ci)[[1]]), ]
beta2<-post_means[grep("^beta.*,2]", names(post_means))]
beta2.ci<-ci[grep("^beta.*,2]", dimnames(ci)[[1]]), ]

gamma<-post_means[grep('gamma', names(post_means))]

gamma.ci<-ci[grep("gamma", dimnames(ci)[[1]]), ]
gamma.ci <-
  cbind.data.frame(gamma,gamma.ci)


theta1<-post_means[grep("theta", names(post_means))]
theta1.ci<-ci[grep("theta", dimnames(ci)[[1]]), ]

delta<-post_means[grep("delta.adj", names(post_means))]
delta.ci<-ci[grep("delta.adj", dimnames(ci)[[1]]), ]
delta.combo<-exp(cbind(delta, delta.ci))

ave.effect.pov<-
  post_means[grep("ave.effect.pov", names(post_means))]
ave.effect.pov.ci<-ci[grep("ave.effect.pov", dimnames(ci)[[1]]), ]
ave.effect.pov.ci <-
  cbind.data.frame(ave.effect.pov,ave.effect.pov.ci)

#RR at last time point
log.rr.vax<-post_means[grep(paste0("^log.rr.vax.*,", ncol(j12.c),"]"), names(post_means)) ]
log.rr.vax.ci<-ci[grep(paste0("^log.rr.vax.*,", ncol(j12.c),"]"), dimnames(ci)[[1]]), ]
log.rr.vax.ci <- cbind.data.frame(log.rr.vax,log.rr.vax.ci)


log.rr.vax.all<-post_means[grep(paste0("log.rr.vax"), names(post_means)) ]
log.rr.vax.all <- 
  t(matrix(log.rr.vax.all, ncol=ncol(j12.c)))

gdp.grp<-rep(1, times=length(gdp.med))
gdp.grp[gdp.med==1]<-2
gdp.grp[gdp.hi==1]<-3

out.list<-list('log.rr.vax'=log.rr.vax,'log.rr.vax.ci'=log.rr.vax.ci,
           'theta1'=theta1,'beta2'=beta2,'beta2.ci'=beta2.ci,'theta1.ci'=theta1.ci,
            'beta1'=beta1,'beta1.ci'=beta1.ci,'mean.vax.c'=mean.vax.c,'sd.vax.c'=sd.vax.c,'ave.effect.pov.ci'=ave.effect.pov.ci,'gdp.grp'=gdp.grp,'posteriors'=posterior_samples.all[seq(from=1,to=sampleN, by=10),])
saveRDS(out.list, 'gdp.output.COL.V1.rds')

```

read in results from pre-run model
```{r}
res1 <- readRDS( 'gdp.output.COL.V1.rds')

ave.effect.pov.ci <- res1$ave.effect.pov.ci
ave.effect.pov.ci.rescale <- ave.effect.pov.ci*(0.9)/sd.vax.c
round(exp(ave.effect.pov.ci.rescale),2)

theta1.ci <-res1$theta1.ci

posteriors <- res1$posteriors
ave.effect.pov.mc<-
  posteriors[,grep("ave.effect.pov", dimnames(posteriors)[[2]] )]
ave.effect.pov.mc.90 <- ave.effect.pov.mc  * 0.9/sd.vax.c
ave.effect.pov.mc.90.hdi <- t(apply(ave.effect.pov.mc.90,2, hdi, credMass = 0.95))
ave.effect.pov.mc.90.med <- apply(ave.effect.pov.mc.90,2, median)
ave.effect.pov.mc.90.hdi <- cbind.data.frame(ave.effect.pov.mc.90.med,ave.effect.pov.mc.90.hdi)
rr_pov_hdi <- exp(ave.effect.pov.mc.90.hdi)
```

### Estimate RR for 90% coverage by mesoregion, separated by GDP
```{r, fig.width=5, fig.height=5}
theta1.rescale <- theta1*(0.9)/sd.vax.c
theta1.ci.rescale <- theta1.ci*(0.9)/sd.vax.c 
exp.ave.ci <- rr_pov_hdi

gdp.grp<-rep(1, times=length(gdp.med))
gdp.grp[gdp.med==1]<-2
gdp.grp[gdp.hi==1]<-3
set.seed(123)
xjit <- rnorm(n=length(gdp.grp),0,0.1)
plot.cols<-c('#e41a1c80','#377eb880','#4daf4a80') #red, blue green
trans.black <- rgb(0,0,0,alpha=0.1)

# plot((gdp.grp+xjit),exp(theta1.rescale),  bty='l', pch=16, ylim=c(0.4,max(exp(theta1.ci.rescale), na.rm=T)),cex=0.75, col=plot.cols[gdp.grp], xlim=c(0.75,3.4), xaxt='n', main='Estimated decline for 90% coverage', xlab='', ylab='Coefficient')

symbols((gdp.grp+xjit),exp(theta1.rescale)  , circles=sqrt(j12.c$`2010` /pi) , bg=plot.cols[gdp.grp],fg='white', inches=0.3,bty='l',ylim=c(0.4,max(exp(theta1.ci.rescale), na.rm=T)),xlim=c(min(gdp.grp)-0.5, max(gdp.grp)+0.5 ), main='Vaccine effect', xlab='',xaxt='n',  col=plot.cols[gdp.grp], xaxt='n', ylab='Rate ratio for 90% uptake')

arrows(x0=(gdp.grp+xjit) , y0=exp(theta1.ci.rescale[,1]) , y1=exp(theta1.ci.rescale[,2]), length=0, col=trans.black)

abline(h=1, col='black', lty=2)

axis(side=1, at=c(1,2,3), labels = c('Low GDP', 'Med GDP', 'High GDP'))
for(i in 1:3){
  x1 <- i + c(0.35,0.4,0.45,0.4,0.35)
  exp.ave.ci <- exp(ave.effect.pov.ci.rescale)
  y1 <- c(exp.ave.ci[i,1],exp.ave.ci[i,3],exp.ave.ci[i,1],exp.ave.ci[i,2],exp.ave.ci[i,1])
  polygon(x1,y1,col=plot.cols[i],border=0)
}
```


### Estimate decline by mesoregion, separated by GDP
slope*uptake in final year
```{r, fig.width=5, fig.height=5}
log.rr.vax.ci <- res1$log.rr.vax.ci
log.rr.vax <- res1$log.rr.vax

gdp.grp<-rep(1, times=length(gdp.med))
gdp.grp[gdp.med==1]<-2
gdp.grp[gdp.hi==1]<-3
set.seed(123)
xjit <- rnorm(n=length(gdp.grp),0,0.1)
plot.cols<-c('#e41a1c80','#377eb880','#4daf4a80') #red, blue green

trans.black <- rgb(0,0,0,alpha=0.1)
# plot((gdp.grp+xjit),exp(log.rr.vax),  bty='l', pch=16, ylim=range(exp(log.rr.vax.ci), na.rm=T), cex=0.75,main='Vaccine effect', xlab='', ylab='Vaccine effect 2015', col=plot.cols[gdp.grp], xaxt='n')

symbols((gdp.grp+xjit),exp(log.rr.vax)  , circles=sqrt(j12.c$`2010` /pi) , bg=plot.cols[gdp.grp],fg='white', inches=0.3,bty='l',ylim=range(exp(log.rr.vax.ci), na.rm=T),main='Vaccine effect', xlab='', ylab='Vaccine effect 2015', col=plot.cols[gdp.grp], xaxt='n')

arrows(x0=(gdp.grp+xjit) , y0=exp(log.rr.vax.ci[,2]) , y1=exp(log.rr.vax.ci[,3]), length=0, col=trans.black)
abline(h=1, col='black', lty=2)

axis(side=1, at=c(1,2,3), labels = c('Low GDP', 'Med GDP', 'High GDP'))

```

Uptake vs RR.
```{r, fig.width=8, fig.height=3}
gdp.grp<-rep(1, times=length(gdp.med))
gdp.grp[gdp.med==1]<-2
gdp.grp[gdp.hi==1]<-3
plot.cols<-c('#e41a1c','#377eb8','#4daf4a') #red, blue green

par(mfrow=c(1,3))
plot( vax.c[,'2011']*100,exp(log.rr.vax.all[7,]), col=plot.cols[gdp.grp], pch=16, bty='l', ylab="Rate Ratio", xlab='3 dose coverage', xlim=c(-50,140))
abline(h=1, lty=2)

plot( vax.c[,'2013']*100,exp(log.rr.vax.all[9,]), col=plot.cols[gdp.grp], pch=16, bty='l', ylab="Rate Ratio", xlab='3 dose coverage', xlim=c(-50,140))
abline(h=1, lty=2)

 plot( vax.c[,'2015']*100,exp(log.rr.vax.all[11,]), col=plot.cols[gdp.grp], pch=16, bty='l', ylab="Rate Ratio", xlab='3 dose coverage', xlim=c(-50,140))
 abline(h=1, lty=2)

```

```{r}
plot.cols<-c('#e41a1c80','#377eb880','#4daf4a80') #red, blue green
widths <- (j12.c$`2010` +j12.c$`2009` +j12.c$`2008` +j12.c$`2007`+j12.c$`2006` +j12.c$`2005` )
widths <- widths/max(widths)*5

plot( vax.c[1,]*100,exp(log.rr.vax.all[,1]), col=plot.cols[gdp.grp[1]], pch=16,type='l', ylim=c(0.6,1.1),bty='l', ylab="Rate Ratio", xlab='3 dose coverage', xlim=c(0,140), lwd=widths[1])

for(i in 1:nrow(vax.c)){
 points( vax.c[i,]*100,exp(log.rr.vax.all[,i]), col=plot.cols[gdp.grp[i]], type='l', bty='l', ylab="Rate Ratio", xlab='3 dose coverage', xlim=c(-50,140), lwd=widths[i])
}
 abline(h=1, lty=2)

```


```{r, fig.width=3, fig.height=3}
plot.cols<-c('#e41a1c','#377eb8','#4daf4a') #red, blue green

 plot( vax.c[,'2015']*100,exp(log.rr.vax.all[11,]), col=plot.cols[gdp.grp], pch=16, bty='l', ylab="Rate Ratio", xlab='3 dose uptake (%)', xlim=c(0,120), ylim=c(0.6, 1.1))
 abline(h=1, lty=2)
```

```{r}
# Time series for 3rd dose coverage by dept.inc
par(mfrow=c(1,1))
matplot(t(vax.c)*100, type="l", bty="l", 
        main="Colombia", axes=F,
        ylab="Coverage (%)", xlab="Year")
axis(2)
axis(side=1,at=1:11,labels=c(2005:2015))
```

# Boxplot for coverage by income level
```{r, fig.width=4, fig.height=5}
par(mfrow=c(1,1))

par(bty='n')
boxplot(vax.c[,'2015']*100  ~ gdp.grp,  ylim=c(0,100), col="grey",
        main="Colombia in 2015",
        names=c("Low","Medium","High"), xlab="Income level", ylab="Coverage (%)")
```